### 2018-07-11 article.Rmd

---
title: "Playing Atari with Keras and R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this article we will implement the famous paper from DeepMind: [*Playing Atari with Deep Reinforcement Learning*](https://www.cs.toronto.edu/~vmnih/docs/dqn.pdf) in R using Keras. This paper introduced the first deep learning model that successfully learned control policies directly from high-dimensional sensory input using reinforcement learning.

The DeepMinds's algorithm is trained on raw pixels from Atari games and estimates future rewards for each possible action. The model is a convolutional neural network trained with a variant of the [Q-Learning](https://en.wikipedia.org/wiki/Q-learning) algorithm, which we will attempt to explain during this article.

The code was not published with the paper, but the results were reproduced and published by severeal people on the internet. Particularly, [this post](https://yanpanlau.github.io/2016/07/10/FlappyBird-Keras.html) and the [referenced implementation](https://github.com/yenchenlin/DeepLearningFlappyBird) were very useful when writing this article. Also, [this article](https://becominghuman.ai/lets-build-an-atari-ai-part-1-dqn-df57e8ff3b26) from [Becoming Human](https://becominghuman.ai) was very inspiring.

# Reinforcement Learning

We recommend reading the [*Part-0: Intro to RL*](https://becominghuman.ai/lets-build-an-atari-ai-part-0-intro-to-rl-9b2c5336e0ec) from Becoming Human as a smooth introduction to [Reinforcement Learning](https://en.wikipedia.org/wiki/Reinforcement_learning).

## Q-Learning

## Deep Q-Learning

# Implementation in R

## Gym